---
title: "renv"
author: "Tom Rushby (@tom_rushby)"
date: "28/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is renv?

An environment manager for R projects. Meaning that it organises the package dependencies within an R project, recording the versions of each package used in analysis.

### Install

```{r cars}
install.packages('renv')
```

Initialise renv and create a project specific local environment with private R library.

```{r}
renv::init()
```

If this is the first use of renv, running the init() command will generate output similar to below:

```Welcome to renv!

It looks like this is your first time using renv. This is a one-time message,
briefly describing some of renv's functionality.

renv maintains a local cache of data on the filesystem, located at:

  - "C:/Users/twr1m15/AppData/Local/R/cache/R/renv"

This path can be customized: please see the documentation in `?renv::paths`.

renv will also write to files within the active project folder, including:

  - A folder 'renv' in the project directory, and
  - A lockfile called 'renv.lock' in the project directory.

In particular, projects using renv will normally use a private, per-project
R library, in which new packages will be installed. This project library is
isolated from other R libraries on your system.

In addition, renv will update files within your project directory, including:

  - .gitignore
  - .Rbuildignore
  - .Rprofile

Please read the introduction vignette with `vignette("renv")` for more information.
You can browse the package documentation online at https://rstudio.github.io/renv/.```

If the project already has a lockfile the following message will be displayed ...

```This project already has a lockfile. What would you like to do? 

1: Restore the project from the lockfile.
2: Discard the lockfile and re-initialize the project.
3: Activate the project without snapshotting or installing any packages.
4: Abort project initialization.```

Refer to [Introduction to renv](https://rstudio.github.io/renv/articles/renv.html) by Kevin Ushey.

### Lock file

The file `renv.lock` contains a description of the state of the project's library.

For example:

{
  "R": {
    "Version": "4.1.1",
    "Repositories": [
      {
        "Name": "CRAN",
        "URL": "https://cran.rstudio.com"
      }
    ]
  },
  "Packages": {
    "renv": {
      "Package": "renv",
      "Version": "0.15.2",
      "Source": "Repository",
      "Repository": "CRAN",
      "Hash": "206c4ef8b7ad6fb1060d69aa7b9dfe69",
      "Requirements": []
    }
  }
}